---
title: "Drag Race Season 11 Episode 5"
author: Thomas Elliott
date: "`r format(Sys.Date(), '%B %e, %Y')`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=8, fig.height=6, fig.retina=4, warning=TRUE)

library(tidyverse)
library(kernlab)
library(e1071)
library(randomForest)
library(neuralnet)
library(ggimage)

season.colors<-c("orange" = "#F7830A",
                 "teal" = "#2CC9B5", 
                 "yellow" = "#FAEE0D",
                 "pink" = "#F60075")

data.df<-read_csv("drag_race.csv")

compareRanks<-function(x, y) {
  #' x = actual place
  #' y = predicted place
  actual<-sum((x-y)^2, na.rm = TRUE)
  worst<-sum((sort(x)-sort(x,TRUE))^2, na.rm = TRUE)
  1-2*(actual/worst)
}

setXY<-function(x,y,n) {
  y = y - round((n-1)/3)*0.8
  if( n %% 3 == 2 ) {
    x<-x - 0.33
  }
  else if ( n %% 3 == 0 ) {
    x<-x + 0.33
  }
  c(x,y)
}

set.seed(11)
```

```{r, include=FALSE}
# generating training and testing sets
all.seasons<-data.df %>% 
  replace_na(list(Wins=0, Highs=0, Lows=0, Lipsyncs=0))
pre11<-all.seasons %>% 
  filter(Season<11)
pre11.factor<-pre11 %>% 
  mutate(Place=factor(Place))
season11<-all.seasons %>% 
  filter(Season==11)
s11.results<-season11 %>% 
  select(Name,filename,Place)
season11<-season11 %>% 
  select(Name,Place,Season,Age,Black,White,POC,PlusSize,Wins,Highs,Lows,Lipsyncs)

the.model<-Place ~ Season + Age + Black + White + POC + PlusSize + Wins + Highs + Lows + Lipsyncs
```

# Episode 5: Monster Ball

[This season](../Episode 01/drag_race_s11e01.md) I re-introduced my machine learning algorithms for predicting the outcomes of Season 11 of RuPaul's Drag Race.

This week established the trend started last season of holding a ball earlier in the season. Previous to season 10, balls were usually held when there were five or four queens left. This season, the ball was held when there were 11 queens left. The theme was Monster Ball, with the queens having to serve three looks, the last of which they had to make themselves in the workroom. The first look was Trampy Trick-or-treater, the second look Witch Please, and the third look MILF: Monsters I'd Like to Freak. This meant 33 total runway looks this week. Overall, I found the ball not too exciting. One of the drawbacks of having the ball earlier is you still have a lot of queens who aren't quite ready to fully compete, and so you get a lot of looks that aren't fully baked. Those were the majority in this episode, with a few stand-outs. Brooke Lynn won the challenge with a mummy trick-or-treater coming down the runway on point (she used to be a professional ballerina), and a very good Witch Doctor look. Her final look was disappointing, though. This was Yvie's challenge to win, and she fell short because of her last look. Her first two, though, were amazing - a T-Rex for her trick-or-treater look and a truly inspired witch look. For her final look, she went for a voodoo doll, but didn't quite nail it. Nina West had great looks for all three themes, but bizarrely she was declared merely safe. Her trick-or-treater was an homage to Little Shop of Horrors, and she brought a unique twist to her witch look as a Salem woman with flames around the hem of her dress. Her final look was truly grotesque. 

In the bottom three were Shuga Cain, for shoddy first and third looks, Ariel Versace for a third look with no monster element, and Silky for disappointing looks all around. Shuga and Ariel have to lip-sync and Ariel sashays away.

Incorporating this week's performances into the data, the algorithms have new predictions for the season:

## Support Vector Machines

```{r}
s11.svm<-ksvm(the.model,
              data=pre11,
              kernal="rbfdot")
predict11<-predict(s11.svm,
                   season11)
s11.results<-s11.results %>% 
  mutate(SVM=rank(predict11,ties.method = "min"))
```


## Gaussian Naive Bayes

```{r}
s11.gnb<-naiveBayes(the.model,
              data=pre11.factor,
              laplace = 0)
predict11<-predict(s11.gnb,
                   season11)
s11.results<-s11.results %>% 
  mutate(GNB=rank(predict11,ties.method = "min"))
```



## Random Forest Classifier

```{r}
s11.rfc<-randomForest(the.model,
              data=pre11.factor,
              ntree=100)
predict11<-predict(s11.rfc,
                   season11)
s11.results<-s11.results %>% 
  mutate(RFC=rank(predict11,ties.method = "min"))
```



## Random Forest Regressor

```{r}
s11.rfr<-randomForest(the.model,
              data=pre11,
              ntree=100)
predict11<-predict(s11.rfr,
                   season11)
s11.results<-s11.results %>% 
  mutate(RFR=rank(predict11,ties.method = "min"))
```



## Neural Networks


```{r}
s11.nn<-neuralnet(the.model,
              data=pre11,
              hidden=3,
              act.fct = "logistic",
              algorithm = "rprop+")
predict11<-predict(s11.nn,
                   season11)
s11.results<-s11.results %>% 
  mutate(NN=rank(predict11,ties.method = "min"))
```



## Final Predictions

The final predicted score is based on the average of predicted places for each algorithm.

The tops and bottoms haven't changed since last week (though there has been some shuffling in the middle). Ra'jah is still predicted to go home next. Brooke Lynn, Yvie, Vanjie, and Silky continue to hold the top four.


```{r}
s11.results<-s11.results %>% 
  mutate(Average=(SVM+GNB+RFC+RFR+NN)/5,
         `Predicted Rank`=rank(Average, ties.method = "min"))
s11.results %>% 
  arrange(Average) %>% 
  select(-filename) %>% 
  knitr::kable()
```

```{r}
s11.score<-compareRanks(s11.results$Place, s11.results[["Predicted Rank"]])
```

Rank Score: `r round(s11.score, digits = 3)`


```{r s11-results, fig.width=10, fig.height=10, fig.retina=4, warning=FALSE, message=FALSE}
df<-s11.results %>% 
  mutate(filename=paste0("../queens/",filename,".jpg")) %>% 
  gather(model,rank,Place:`Predicted Rank`) %>% 
  filter(model!="Average") %>% 
  mutate(model=factor(model, levels = c("Place", "SVM", "GNB", "RFC", "RFR", "NN", "Predicted Rank")),
         x=as.numeric(model),
         y=16-rank)
labels<-df %>% 
  filter(model=="Predicted Rank") %>% 
  mutate(x=8)
for(i in unique(df$model) ) {
  places<-list()
  for( n in unique(df$Name) ) {
    temp<-df %>% filter(Name==n&model==i)
    this.rank<-as.character(temp$rank)
    if( !is.na(this.rank) ) {
      if( this.rank %in% names(places) ) {
        places[[this.rank]]<-places[[this.rank]]+1
      } else {
        places[[this.rank]]<-1
      }
      new.xy<-setXY(temp$x,temp$y,places[[this.rank]])
      df$x[df$Name==n&df$model==i]<-new.xy[1]
      df$y[df$Name==n&df$model==i]<-new.xy[2]
    }
  }
}
df %>% 
  ggplot(aes(x=x, y=y)) +
  geom_image(aes(image=filename)) +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.grid.major.y = element_line(color=season.colors),
        axis.text.y = element_text(color="black", size=12),
        axis.ticks.y = element_line(color=season.colors)) +
  scale_x_continuous(name=NULL, breaks=c(1:7), 
                     labels=c("Actual Place", "SVM", "GNB", "RFC", "RFR", "NN", "Predicted Place")) +
  scale_y_continuous(name="Place", breaks=c(1:15), labels=c(15:1)) +
  labs(title="RuPaul's Drag Race Season 11")
ggsave("season11_results.png", width=10, height=10, dpi=400)
```

